---
title: "Lab1"
author: "Eleanore Brugh"
date: "2026-02-22"
output: html_document
---

###lab exercises
```{r}
library(tidyverse)
library(ggplot2)
library(sf)
library(dplyr)
```

```{r}
p.counties <- "C:/R/Data/CBW/County_Boundaries.shp"
p.stations <- "C:/R/Data/CBW/Non-Tidal_Water_Quality_Monitoring_Stations_in_the_Chesapeake_Bay.shp"
```

```{r}
d.counties <- sf::read_sf(p.counties)
d.stations <- sf::read_sf(p.stations)
```

```{r}
glimpse(d.counties)
glimpse(d.stations)
```

```{r}
d.stations %>% sf::st_is_valid()
```

```{r}
d.counties %>% sf::st_is_valid()
```

```{r}
d.counties <- d.counties %>% sf::st_make_valid()
```


```{r}
d.counties %>% dplyr::select(GEOID10, ALAND10) %>% head()
```

```{r}
d.counties %>% dplyr::select(-NAME10) %>% head() 
```

```{r}
d.counties %>% dplyr::select(GEOID10:CLASSFP10) %>% head() 
```

```{r}
d.counties %>% dplyr::select(-(GEOID10:CLASSFP10)) %>% head()
```

```{r}
d.counties %>% dplyr::select(starts_with("C"))
```

#grouping data
```{r}
d.counties %>% group_by(STATEFP10) %>% mutate(stateLandArea = sum(ALAND10))
```

#without geometry
```{r}
d.counties %>% {
  as_tibble(d.counties) %>% dplyr::select(-geometry) %>%
    group_by(STATEFP10) %>%
    summarise(stateLandArea = sum(ALAND10))}
```

#plot (+)
```{r}
d.counties %>% {
  ggplot(., aes(x = as.factor(STATEFP10), y = ALAND10)) +
    geom_boxplot(aes(fill = STATEFP10))}
```

```{r}
d.counties %>% {
  ggplot(., aes(x = ALAND10)) +
    geom_histogram(aes(fill = STATEFP10)) +
    labs(title = "not useful")}
```

##spatial operations
#same CRS
```{r}
d.counties %>% sf::st_crs() == d.stations %>% sf::st_crs()
```

#trimming data
```{r}
del.counties <- d.counties %>% dplyr::filter(STATEFP10 == 10)
```

#spatial intersection
```{r}
del.stations <- sf::st_intersection(d.stations, del.counties)
```

```{r}
glimpse(del.stations)
```
```{r}
plot(del.stations)
```
```{r}

```
```{r}
del.counties %>% st_area()
```






#TASK 1
####Task1-1
```{r}
task1_1 <- d.counties %>% 
  as_tibble() %>% dplyr::select(-geometry) %>%
  group_by(STATEFP10) %>%
  mutate(totalStateLandArea = sum(ALAND10 + AWATER10),
  (land_percent_total_area = ALAND10 / totalStateLandArea) * 100) 
```

####Task1-2
######calculated the water proportion group function not needed
```{r}
prop_calc <- d.counties %>%
  as_tibble() %>% dplyr::select(-geometry) %>%
  mutate(water_proportion = AWATER10 / (ALAND10 + AWATER10))
```

```{r}
glimpse(prop_calc)
```

#####find max value and create new table
#####need state, county, and water proportion
```{r}
task1_2 <- prop_calc %>%
  group_by(STATEFP10) %>%
  slice_max(water_proportion, n = 1) %>%
  select(STATEFP10, NAME10, water_proportion)
```

```{r}
glimpse(task1_2)
```

####Task1-3
#####no group just calculation
```{r}
task1_3 <- d.counties %>%
  as_tibble() %>% dplyr::select(-geometry) %>%
  count(STATEFP10, name = "count_perstate")
```

```{r}
glimpse(task1_3)
```
####Task1-4
#####calculate characters
```{r}
name_length <- d.stations %>%
  as_tibble() %>% dplyr::select(-geometry) %>%
  mutate(tot_char = nchar(STATION_NA))
```

```{r}
glimpse(name_length)
```

#####find shortest name
```{r}
task1_4 <- name_length %>%
  slice_min(name_length, n = 1) %>%
  select(STATION_NA, tot_char)
```
####answer for 1-4
```{r}
glimpse(task1_4)
```

#TASK 2
####Task2-1
#####scatterplot
```{r}
ggplot(d.counties) +
  aes(x = ALAND10, y = AWATER10, color = STATEFP10) + geom_point(size = 1) +
  labs(title = "Relationship Bewteen Land and Water Area for Counties", x = "Land Area (ALAND10)", y = "Water Area (AWATER10)")
```
####Task2-2
#####histogram
```{r}
ggplot(d.stations) +
  aes(x = Drainage_A) +
  geom_histogram(bins = 15) +
  labs(title = "Drainage Area for All Monitoring Stations", x = "Drainage Area (Drainage_A)", y = "Number of Stations")
```

####Task2-3
#####join - not intersection
```{r}
join_stations_state <- st_join(d.stations, d.counties["STATEFP10"])
```
#####histogram
```{r}
ggplot(join_stations_state) +
  aes(x = Drainage_A, fill = STATEFP10) +
  geom_histogram(bins = 30) +
  labs(title = "Drainage Area for All Monitoring Stations", x = "Drainage Area (Drainage_A)", y = "Number of Stations", fill = "STATEFP10")
```

#Task 3
####Task3-1
```{r}
f.efunc <- function(x){
  if (!is.numeric(x)) stop("Error") 
    mean = mean(x)
    median = median(x)
    maximum = maximum(x)
    minimum = minimum(x)
    sort(x)}
```









